### 2014-11-13 doc/thesis/system.tex
Updating typing rules
Minor updates to the typing rules sketch

\begin{figure}[!ht]
\textbf{Abstract Syntax}\\
\dstart
$$
\begin{array}{rl}
s ::= & \;\; \mathbf{skip} \; | \;
s_{1} \; ; \; s_{2} \; | \;
\vec{l} = el  \; | \;
\mathbf{while} \; e \; \mathbf{do} \; s \; | \;
\mathbf{if} \; e \; \mathbf{then} \; s_{1} \; \mathbf{else} \; s_{2}\\
& | \; \mathbf{local} \; \vec{n{:}t} = el \; \mathbf{in} \; s \; | \;
\mathbf{rec} \; n{:}t = f \; \mathbf{in} \; s \; | \;
\mathbf{return} \; el \; | \;
a\\
e ::= & \;\; \mathbf{nil} \; | \;
\mathbf{false} \; | \;
\mathbf{true} \; | \;
{\it int} \; | \;
{\it float} \; | \;
{\it string} \; | \;
{...} \; | \;
f \; | \;
\{ \; cl \; \} \\
& | \; e_{1} + e_{2} \; | \;
e_{1} \; {..} \; e_{2} \; | \;
e_{1} == e_{2} \; | \;
e_{1} < e_{2} \; | \;
e_{1} \; \mathbf{and} \; e_{2} \; | \;
e_{1} \; \mathbf{or} \; e_{2} \\
& | \; \mathbf{not} \; e \; | \;
- e \; | \;
\# \; e \; | \;
a \; | \;
l \; | \;
{<}t{>} \; n\\
l ::= & \;\; n \; | \;
e_{1}[e_{2}]\\
el ::= & \;\; \mathbf{nothing} \; | \;
\vec{e} \; | \;
me \; | \;
\vec{e}, me\\
cl ::= & \;\; \mathbf{nothing} \; | \;
\vec{c} \; | \;
\vec{e} \; | \;
me \; | \;
\vec{c}, \vec{e} \; | \;
\vec{c}, me \; | \;
\vec{c}, \vec{e}, me \; | \;
\vec{e}, me\\
me ::= & \;\; a \; | \;
{...}\\
a ::= & \;\; e_{1}(\vec{e_{2}}) \; | \;
e_{1}{:}n(\vec{e_{2}})\\
f ::= & \;\; \mathbf{fun} \; (){:}r \; s \; ; \; \mathbf{return} \; el \; | \;
\mathbf{fun} \; ({...}{:}t){:}r \; s \; ; \; \mathbf{return} \; el\\
& | \; \mathbf{fun} \; (\vec{n{:}t}){:}r \; s \; ; \; \mathbf{return} \; el \; | \;
\mathbf{fun} \; (\vec{n{:}t},{...}{:}t){:}r \; s \; ; \; \mathbf{return} \; el\\
c ::= & \;\; [e_{1}] = e_{2}\\
n ::= & \;\; {\it name}\\
\end{array}
$$
\dend
\caption{Typed Lua abstract syntax}
\label{fig:syntax}
\end{figure}

@@

\begin{verbatim}
    rec f:(number) -> (number) =
      fun (x:number)
        if n == 0 then
          return 1
        else
          return n * f(n - 1)
    in f(5)
\end{verbatim}

@@

\begin{verbatim}
    local a:(number) -> (number) =
      fun (x:number)
        if n < 0 then return -n else return n
    in local d:(any, any) -> (number) =
         fun (x:any, y:any)
           skip ; return a(x - y)
       in d(1, 1)
\end{verbatim}

@@

\begin{verbatim}
  local m:() -> (number, string) =
    fun () skip ; return 2, "foo"
  in local s:(number, number) -> (number)
       fun (x:number, y:number) skip ; return x + y
     in local x:number, y:number, z:string = m(), m()
        in s(m(), m())
\end{verbatim}

@@

\begin{verbatim}
    local t:{string:number|nil} = { ["foo"] = 1 } in
      local x:number|nil = t["foo"] in x = t["bar"]
\end{verbatim}

@@

\begin{verbatim}
    local d:{integer:string|nil} =
      { "Sun", "Mon", "Tue", "Wed", "Thu", "Fri", "Sat" }
    in local x:number|nil = d[1] in x = d[8]
\end{verbatim}

@@

\begin{verbatim}
    local s:{"x":number, "y":number} = { x = 0, y = 0 }
    in s["new"] <(self, number, number) -> (self)> =
       fun (self:self, x:number, y:number):(self)
         local s:{"x":number, "y":number} =
           setmetatable({}, { __index = self })
         in s["x"] = x ; s["y"] = y ; return s
     ; s["move"] <(self, number, number) -> ()> =
       fun (self:self, x:number, y:number):()
         self["x"] = self["x"] + x ;
         self["y"] = self["y"] + y 
\end{verbatim}

>>>>

\begin{figure}[!ht]
\textbf{Abstract Syntax}\\
\dstart
$$
\begin{array}{rl}
s ::= & \;\; \mathbf{skip} \; | \;
s_{1} \; ; \; s_{2} \; | \;
\vec{l} = el \; | \;
\mathbf{while} \; e \; \mathbf{do} \; s \; | \;
\mathbf{if} \; e \; \mathbf{then} \; s_{1} \; \mathbf{else} \; s_{2}\\
& | \; \mathbf{local} \; \vec{n{:}t} = el \; \mathbf{in} \; s \; | \;
\mathbf{rec} \; n{:}t = f \; \mathbf{in} \; s \; | \;
\mathbf{return} \; el \; | \;
e(el)_{s} \; | \;
e{:}n(el)_{s}\\
e ::= & \;\; \mathbf{nil} \; | \;
\mathbf{false} \; | \;
\mathbf{true} \; | \;
{\it int} \; | \;
{\it float} \; | \;
{\it string} \; | \;
{...}_{e} \; | \;
f \; | \;
\{ \; cl \; \} \\
& | \; e_{1} + e_{2} \; | \;
e_{1} \; {..} \; e_{2} \; | \;
e_{1} == e_{2} \; | \;
e_{1} < e_{2} \; | \;
e_{1} \; \mathbf{and} \; e_{2} \; | \;
e_{1} \; \mathbf{or} \; e_{2} \\
& | \; \mathbf{not} \; e \; | \;
- e \; | \;
\# \; e \; | \;
e(el)_{e} \; | \;
e{:}n(el)_{e} \; | \;
n \; | \;
e_{1}[e_{2}] \; | \;
{<}t{>} \; n\\
l ::= & \;\; n \; | \;
e_{1}[e_{2}] \; | \;
n[e] \; {<}t{>} \; | \;
n_{1}{:}n_{2} \; {<}t{>}\\
el ::= & \;\; \mathbf{nothing} \; | \;
\vec{e} \; | \;
me \; | \;
\vec{e}, me\\
cl ::= & \;\; \mathbf{nothing} \; | \;
\vec{c} \; | \;
\vec{e} \; | \;
me \; | \;
\vec{c}, \vec{e} \; | \;
\vec{c}, me \; | \;
\vec{c}, \vec{e}, me \; | \;
\vec{e}, me\\
me ::= & \;\; e(el) \; | \;
e{:}n(el) \; | \;
{...}\\
f ::= & \;\; \mathbf{fun} \; (){:}r \; s \; ; \; \mathbf{return} \; el \; | \;
\mathbf{fun} \; ({...}{:}t){:}r \; s \; ; \; \mathbf{return} \; el\\
& | \; \mathbf{fun} \; (\vec{n{:}t}){:}r \; s \; ; \; \mathbf{return} \; el \; | \;
\mathbf{fun} \; (\vec{n{:}t},{...}{:}t){:}r \; s \; ; \; \mathbf{return} \; el\\
c ::= & \;\; [e_{1}] = e_{2}\\
n ::= & \;\; {\it name}\\
\end{array}
$$
\dend
\caption{Typed Lua abstract syntax}
\label{fig:syntax}
\end{figure}

@@

\begin{verbatim}
    rec f:(integer) -> (integer) =
      fun (x:integer)
        if n == 0 then
          return 1
        else
          return n * f(n - 1)
    in f(5)
\end{verbatim}

@@

\begin{verbatim}
    local a:(integer) -> (integer) =
      fun (x:integer)
        if n < 0 then return -n else return n
    in local d:(any, any) -> (integer) =
         fun (x:any, y:any)
           skip ; return a(x - y)
       in d(1, 1)
\end{verbatim}

@@

\begin{verbatim}
  local m:() -> (integer, string) =
    fun () skip ; return 2, "foo"
  in local s:(integer, integer) -> (integer)
       fun (x:integer, y:integer) skip ; return x + y
     in local x:integer, y:integer, z:string = m(), m()
        in s(m(), m())
\end{verbatim}

@@

\begin{verbatim}
    local t:{string:integer|nil} = { ["foo"] = 1 } in
      local x:integer|nil = t["foo"] in x = t["bar"]
\end{verbatim}

@@

\begin{verbatim}
    local d:{integer:string|nil} =
      { "Sun", "Mon", "Tue", "Wed", "Thu", "Fri", "Sat" }
    in local x:string|nil = d[1] in x = d[8]
\end{verbatim}

@@

\begin{verbatim}
    local s:{"x":number, "y":number} = { x = 0., y = 0.0 }
    in s["new"] <(self, number, number) -> (self)> =
       fun (self:self, x:number, y:number):(self)
         local s:{"x":number, "y":number} =
           setmetatable({}, { __index = self })
         in s["x"] = x ; s["y"] = y ; return s
     ; s["move"] <(self, number, number) -> ()> =
       fun (self:self, x:number, y:number):()
         self["x"] = self["x"] + x ;
         self["y"] = self["y"] + y 
\end{verbatim}
